Data Exploration

### 2018-11-27 src/data_exploration.Rmd
background and presence correlation plots and output

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
The "include=FALSE" supresses the package installation text appearing in the document...
```{r pre-install packages, include=FALSE}
packages <- c("plyr", "graphics", "ggplot2") 
source("../src/ipak.R")
ipak(packages)
```

@@

>>>>

# Package dependencies
You can load them using the following code which uses a function called [ipak](https://gist.github.com/stevenworthington/3178163). 
Note this function checks to see if the packages are installed first.
The "include=FALSE" supresses the package installation text appearing in the document...
```{r pre-install packages, include=FALSE}
packages <- c("plyr", "graphics", "ggplot2", "corrplot") 
source("../src/ipak.R")
ipak(packages)
```

@@

## correlations between background points

check to see if there are any correlations in the env. variables

first subset the env.correlate columns (you don't need everything)

```{r}
background_env <- subset(background, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
background_env_cor <- cor(background_env, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(background_env_cor, "../output/env/background_env_corr.csv", row.names = TRUE)
background_corrplot <- corrplot(background_env_cor , method = "color", type = "upper", order = "alphabet", tl.cex = 0.8) 
dev.copy(png,"../output/env/simple_plots/background_envcorr.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

## correlations between presence points

check to see if there are any correlations in the env. variables

first subset the env.correlate columns (you don't need everything) then use cor to get the correlation values, and then corrplot for a visual

```{r}
pres_env <- subset(presence, select = c(amo_sample, amo_winter, amo_prev, nao_sample, nao_winter, nao_prev, chl_surface, chl_depth, mlp_surface, ssh_surface, temp_surface, temp_depth, o2_surface, o2_depth, salinity_surface, salinity_depth))
pres_env_cor <- cor(pres_env, use="complete.obs", method = "spearman") #need to create a correlation using cor to plot. complete.obs means na.rm in this package. #using spearmans as suitable for non-linear relationships
write.csv(pres_env_cor, "../output/env/pres_env_corr.csv", row.names = TRUE)
pres_corrplot <- corrplot(pres_env_cor , method = "color", type = "upper", order = "alphabet", tl.cex = 0.8) 
dev.copy(png,"../output/env/simple_plots/pres_envcorr.png") # to automatically save the plot to a png AND show it inline
dev.off() # stops automatic saving of the plot to a png
```

```{r}
pairs.panels(pres_env_cor)
```
